Supplementary Report for the article *Functional diversity can facilitate the collapse of an undesirable ecosystem state*

### 2022-11-24 reports/supplement/supplement_0.9.1.Rmd

Research about the relationship between biodiversity and ecosystem functioning has identified two major mechanisms by which diversity affects ecosystem functioning: the selection/sampling effect and the complementarity effect (Loreau & Hector 2001). The sampling effect occurs when more diverse communities have higher functioning because they have a higher probability of containing the best-performing species, and when only one (or few) of the species in a community determine the community and ecosystem level properties, such as biomass production or stability (Cardinale et al. 2011). In contrast, the complementarity effect occurs when higher ecosystem functioning in more diverse communities is due to niche differences or positive interactions among species, i.e. the diversity effect is driven by combinations of the species in a community (Cardinale et al. 2011). Both effects are considered valid and important mechanisms by which diversity affects ecosystem functioning (Cardinale et al. 2006, 2011, 2012; Fargione et al. 2007).

>>>>

Research about the relationship between biodiversity and ecosystem functioning has identified two major mechanisms by which diversity affects ecosystem functioning: the selection/sampling effect and the complementarity effect (Loreau & Hector 2001). The sampling effect occurs when more diverse communities have higher functioning because they have a higher probability of containing a high performing species, and when only one (or few) of the species in a community determine the community and ecosystem level properties, such as biomass production or stability. In this case the observed effect of diversity is driven by only one (or few) species (Cardinale et al. 2011).  In contrast, the complementarity effect occurs when higher ecosystem functioning in more diverse communities is due to niche differences or positive interactions among species, i.e. the diversity effect is driven by combinations of the species in a community (Cardinale et al. 2011). Both effects are considered valid and important mechanisms by which diversity affects ecosystem functioning (Cardinale et al. 2006, 2011, 2012; Fargione et al. 2007).
